---
title: "RLadies Meetup Groups - No First Meetup"
author: "Augustina Ragwitz"
date: "January 22, 2018"
output: html_document
params:
  meetup_api_key: !r Sys.getenv("API_KEY_MEETUP_RLADIES")
  archive_folder: !r Sys.Date()
  output_folder: "latest"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(ggthemes)
library(httr)
library(jsonlite)
library(lubridate)
library(maps)
library(purrr)
library(readr)
library(reshape2)
library(scales)
library(stringr)
library(tidyr)
```

# Meetup Data

```{r globals, warning=FALSE}
dir.create(file.path(params$output_folder))
dir.create(file.path(params$archive_folder))

get_meetups <- function (url, query) {
  req <- GET(url, query=query)
  print(paste(req$url))
  json <- content(req, as = "text")
  things <- fromJSON(json, flatten=TRUE)
  return(things)
}
```

## R-Ladies Meetup Groups


```{r find_groups}

get_rladies_groups <- function (folder) {
  
  groups_url <- "https://api.meetup.com/find/groups"

  groups_query_params <- list(
    key=params$meetup_api_key, 
    sign=TRUE,
    page=200,
    radius="global")
  
  # by topic
  meetup_groups_topic <- get_meetups(groups_url, append(groups_query_params, c(topic_id=1513883, order="members")))
  
  # by text + category
  meetup_groups_text <- get_meetups(groups_url, append(groups_query_params, c(text="r-ladies", category=34)))
  
  meetup_groups_aggr <- bind_rows(meetup_groups_topic, meetup_groups_text %>% anti_join(meetup_groups_topic, by="id"))
  meetup_groups <- meetup_groups_aggr %>% filter(str_detect(name, "[Rr]([ -]?)[Ll]adies"))
  
  meetup_groups <- meetup_groups %>%
    mutate(events_url = paste("https://api.meetup.com/", urlname, "/events", sep="")) %>%
    select(-meta_category.category_ids) # remove lists
  
  write_csv(meetup_groups, paste(folder, "rladies_meetup_groups.csv", sep="/"), 
            na = "")
  write_csv(meetup_groups_topic %>% select(-meta_category.category_ids), 
            paste(folder, "rladies_meetup_groups_topic.csv", sep="/"), 
            na = "")
  write_csv(meetup_groups_text %>% select(-meta_category.category_ids), 
            paste(folder, "rladies_meetup_groups_text.csv", sep="/"), 
            na = "")
  return(meetup_groups)
}

```

## R-Ladies Meetup Events

Filter out groups with no next event.

```{r get_no_upcoming, message=FALSE}
meetup_groups <- read_csv(paste(params$output_folder, "rladies_meetup_groups.csv", sep="/"))

get_rladies_groups_no_upcoming <- function(folder, groups){
  meetup_groups_no_upcoming <- meetup_groups %>% filter(is.na(next_event.id))
  write_csv(meetup_groups_no_upcoming, 
            paste(folder, "rladies_meetup_groups_no_upcoming.csv", sep="/"), 
            na = "")
  return(meetup_groups_no_upcoming)
}

```

For groups that don't have a next event, get the last event they had.

```{r get_events, eval=FALSE}
# This gets throttled, can't use function

#get_rladies_recent_past_events <- function (folder, meetup_groups) {
meetup_groups_no_upcoming <- no_upcoming
folder <- params$archive_folder

  recent_past_events_query_params <- list(
    key=params$meetup_api_key, 
    sign=TRUE,
    status="past",
    scroll="recent_past"
  )
  
  dir.create(file.path(paste(folder, "rladies_meetup_events_recent_past", sep="/")))
  
  meetup_events_past <- data_frame()

  for (n in 1:nrow(meetup_groups_no_upcoming)) {
    events_url <- meetup_groups_no_upcoming[n,]["events_url"]
    print(paste("Trying url:", events_url$events_url))
    meetups <- get_meetups(events_url$events_url, recent_past_events_query_params)
    if(length(meetups)== 0) {
      next()
    }
  
    meetups <- meetups %>% mutate(photo_album.photo_sample=NA)
    meetup_events_past <- bind_rows(meetup_events_past, meetups)
    write_csv(meetup_events_past, paste0(folder, "/rladies_meetup_events_recent_past/_meetup_events_recent_past_", n, ".csv"),
              na = "")
  }
  
  write_csv(meetup_events_past, paste(folder,"rladies_meetup_events_recent_past.csv", sep="/"), 
            na = "")
  #return(meetup_events_past)
#}
```

## Get Latest

Create a new archive folder and get the latest data.

```{r get_latest_data, eval=FALSE, message=FALSE}

# Run this first when updating archives
dir.create(file.path(params$archive_folder))

meetup_groups <- get_rladies_groups(params$archive_folder)
groups_no_upcoming <- get_rladies_groups_no_upcoming(params$archive_folder, meetup_groups)

# TODO add try/catch for throttling - or just pull past events since last pull and merge with last pull
# past_events <- get_rladies_past_events(params$archive_folder, groups_no_upcoming)
# Run ^^ manually in above code block for now then run code block below to copy archive folder to output

```

```{r copy_archive_to_latest, eval=FALSE}

system2("rm", args = c("-r", params$output_folder))
system2("cp", args = c("-R", paste(params$archive_folder), paste(params$output_folder)))

```

# Meetup Group Analysis

## Meetup Groups with No First Event

Groups that haven't yet had a first event. Organizers should be contacted to see how we can help them.

```{r no_first_event, message=FALSE}

no_upcoming <- read_csv(paste(params$output_folder, "rladies_meetup_groups_no_upcoming.csv", sep="/"))
past_events <- read_csv(paste(params$output_folder, "rladies_meetup_events_recent_past.csv", sep="/"))

no_first_event <- no_upcoming %>% 
  anti_join(past_events %>% select(group.id), by=c("id"="group.id"))

no_first_event <- no_first_event %>%
  mutate(created=as.POSIXct(created/1000, tz="UTC", origin="1970-01-01"),
         meetup_age=difftime(Sys.Date(), created, unit="days"),
         chapter=str_replace(name, "R[ -]?[Ll]adies( )?(- )?", "")) %>%
  separate(timezone, c("region"), extra="drop", fill="left")
         

```

```{r no_first_event_plot, message=FALSE, fig.width=10, fig.height=8}
       
ggplot(no_first_event, aes(x=reorder(chapter, meetup_age), y=meetup_age)) +
  geom_bar(aes(fill=region), stat="identity") +
  theme_few() +
  coord_flip() +
  labs(x="Chapter", y="Age (Days)", title="Meetup Groups with no First Event")
```

## Meetup Groups with Event Lag

Groups that haven't had an event in awhile. Organizers should be contacted to see how we can help them.

```{r event_lag, message=FALSE}

no_upcoming <- read_csv(paste(params$output_folder, "rladies_meetup_groups.csv", sep="/"))
past_events <- read_csv(paste(params$output_folder, "rladies_meetup_events_recent_past.csv", sep="/"))

latest_events <- past_events %>% 
  filter(status=="past") %>%
  rename(event.id = id, event.time = time) %>%
  group_by(group.name) %>%
  top_n(1, event.time)

groups_lagged <- no_upcoming %>%
    inner_join(latest_events %>% 
                 select(group.id, event.id, event.time), 
               by=c("id"="group.id"))

groups_lagged <- groups_lagged %>%
  rename(last_event=event.time) %>%
  mutate(created=as.POSIXct(created/1000, tz="UTC", origin="1970-01-01"),
         meetup_age=difftime(Sys.Date(), created, unit="days"),
         chapter=str_replace(name, "R[ -]?[Ll]adies( )?(- )?", ""),
         last_event=as.POSIXct(last_event/1000, tz="UTC", origin="1970-01-01"),
         last_event_days=difftime(Sys.Date(), last_event, unit="days")
         ) %>%
  separate(timezone, c("region"), extra="drop", fill="left")

# Special case chapter names
groups_lagged <- groups_lagged %>% 
  mutate(chapter=ifelse(chapter=="Spotkania Entuzjastów R-Warsaw RUG Meetup & Warsaw", "Warsaw", chapter))

groups_lagged %>% select(chapter, last_event_days, last_event) %>% arrange(-last_event_days)

```

```{r groups_no_next, message=FALSE, fig.width=10, fig.height=8}

ggplot(groups_lagged, aes(x=reorder(chapter, last_event_days), y=last_event_days)) +
  geom_bar(aes(fill=region), stat="identity") +
  theme_few() +
  coord_flip() +
  labs(x="Chapter", y="Last Event (Days)", title="Meetup Groups with no Next Event")

```

```{r groups_lagged, message=FALSE, fig.width=10, fig.height=8}


ggplot(groups_lagged %>% filter(last_event_days > 60), aes(x=reorder(chapter, last_event_days), y=last_event_days)) +
  geom_bar(aes(fill=region), stat="identity") +
  theme_few() +
  coord_flip() +
  labs(x="Chapter", y="Last Event (Days)", title="Meetup Groups with no Next Event (> 60 days since last event)")

```



## Meetup Groups Not Using R-Ladies Topic

Groups that aren't showing up the R-Ladies Topic search. Organizers should add the R-Ladies topic to their group description.

```{r no_topic, message=FALSE}

meetup_groups_topic <- read_csv(paste(params$output_folder, "rladies_meetup_groups_topic.csv", sep="/"))
meetup_groups_text <- read_csv(paste(params$output_folder, "rladies_meetup_groups_text.csv", sep="/"))

# group + organizer to contact
meetups_missing_topic <- meetup_groups_text %>% anti_join(meetup_groups_topic, by="id")

# Special case chapter names
meetups_missing_topic <- meetups_missing_topic %>% 
  mutate(name=ifelse(name=="Spotkania Entuzjastów R-Warsaw RUG Meetup & R-Ladies Warsaw", "R-Ladies Warsaw", name))

meetups_missing_topic %>% filter(str_detect(name, "[Rr]([ -]?)[Ll]adies")) %>%
  arrange(-members) %>% 
  select(name, localized_location, organizer.name, members, created)

```




